---
title: "hw6ChetnaKhanna"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(maps)
library(RColorBrewer)
```

```{r}
#loading the species csv
df_species <- read.csv('species.csv')
head(df_species)
```

```{r}
#loading the parks csv
df_parks <- read.csv('parks.csv')
head(df_parks)
```

```{r}
#left joining of the two tables
df <- merge(x = df_parks, y = df_species, by = "Park.Name", all.x = TRUE)
head(df)
#summary(df)
```

We can see that 114530 species have null values in the Conservation.Status. By looking at this number, we can say that these species must be of least concern. So let us replace null by Least Concern in Conservation.Status column.

```{r}
#let us replace NA in Conservation.Status to 'Safe'
con_status <- function(x) {
  if (x == '') {return ('Least Concern')}
  else {return (as.character(x))}
}

#using sapply function to replace empty status with Safe
df$Conservation.Status <- sapply(df$Conservation.Status, con_status)

#looking at the dataframe once again
head(df)
```

Let us have a look at the different Categories of species.
```{r}
#let us see the count of different Categories of species
table(df$Category)
```

We can see that the data frame have the following species of plants:
1. Algae
2. Fungi
3. Nonvascular Plant
4. Vascular Plant

We can see that the data frame have the following species of animals:
1. Amphibian
2. Bird
3. Crab/Lobster/Shrimp
4. Fish
5. Insect
6. Invertebrate
7. Mammal
8. Reptile
9. Slug/Snail 
10. Spider/Scorpion

Let us add a new column 'Main.Category' to our data frame with two values - Animal and Plant.

```{r}
#adding a new column Main.Category with value Animal or Plant
df$Main.Category <- ifelse (df$Category %in% c('Algae','Fungi','Nonvascular Plant', 'Vascular Plant'), 'Plant', 'Animal')
#changing the character type to factor type to get the levels
df$Main.Category <- as.factor(df$Main.Category)
df$Conservation.Status <- as.factor(df$Conservation.Status)
summary(df)
#head(df)
```

We can see that Park.Name and Park.Code variable has many repeated enteries as one park can have many species. So, let us group Park.Code and get the count of each park.

```{r}
#let us see the different Parks 
table(df$Park.Code)
```

```{r}
df <- df %>% group_by(Park.Code) %>% mutate(Total = n())
head(df)
```

Let us draw the map depicting the species count in US.

```{r}
df %>% ggplot(aes(x = Longitude, y = Latitude)) +
  borders("state") +
  geom_point(aes(size=Total, color=Total)) +
  coord_map() + 
  labs(title="US National Parks", subtitle="Species Biodiversity") +
  scale_colour_distiller(palette="Oranges") +
  labs(size="Species Count", colour="Species Count")
```

Due to Alaska and Hawaii, the map dots look clutered. Let us try to use some other way to depict Alaska and Hawaii. 
```{r}
df_cord <- data.frame(df$Longitude, df$Latitude)
df_map <- usmap_transform(df_cord)

plot_usmap() + 
  geom_point(data=df_map, aes(x=df.Longitude.1, y=df.Latitude.1), color="red") +
  labs(title="US National Parks", subtitle="Species Biodiversity") +
  scale_colour_distiller(palette="Oranges") +
  labs(size="Species Count", colour="Species Count")
```

Let us make the above graph visually appealing. 
```{r}
df_cord <- data.frame(df$Longitude, df$Latitude)
df_trans <- usmap_transform(df_cord)
df_map <- merge(df, df_trans, by.x="Latitude", by.y="df.Latitude")

plot_usmap(fill="cornsilk2") + 
  geom_point(data=df_map, aes(x=df.Longitude.1, y=df.Latitude.1, size=Total, color=Total)) +
  labs(title="US National Parks", subtitle="Species Count") +
  scale_colour_distiller(palette="Oranges") +
  labs(size="Species Count", colour="Species Count") + 
  theme(legend.position="right", plot.title=element_text(face="bold", size=16), plot.subtitle=element_text(face="bold", size=13))
```

Let us find the count of national parks in different states of US.

```{r}
df_parks %>% group_by(State) %>% summarise(n=n()) %>%
  ggplot() +
  geom_bar(aes(reorder(x=State, -n), y=n), stat="identity", fill="dark green") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  labs(x="State",y="Count", title="National Parks in US", subtitle="State-wise Count") + 
  theme(axis.title.x=element_text(face="bold", size=12), 
        axis.title.y=element_text(face="bold", size=12), 
        plot.title=element_text(face="bold", size=16), 
        plot.subtitle=element_text(face="bold", size=13), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.line=element_line(colour = "black"))
```

Let us find the top 10 US states with maximum number of national parks.

```{r}
count_park <- df_parks %>% group_by(State) %>% summarise(n=n())

head(arrange(count_park, desc(n)), n=10) %>%
  ggplot() +
  geom_bar(aes(reorder(x=State, -n), y=n), stat="identity", fill="dark green") + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  labs(x="State",y="Count", title="National Parks in US", subtitle="State-wise Count") + 
  theme(axis.title.x=element_text(face="bold", size=12), 
        axis.title.y=element_text(face="bold", size=12), 
        plot.title=element_text(face="bold", size=16), 
        plot.subtitle=element_text(face="bold", size=13), 
        panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background=element_blank(), 
        axis.line=element_line(colour = "black")) +
  [scale_fill_brewer(palette="OrRd")
```

```{r}
table(df$Conservation.Status)
```

Let us have a look at the species which are under concern. So we will use only the following categories for our graph:
1. Endangered 
2. In Recovery 
3. Proposed Endangered
4. Proposed Threatened
5. Species of Concern
6. Threatened
7. Under Review
Also, let us have a look at only animal species.

```{r}
df %>% filter(Conservation.Status == c("Endangered","In Recovery","Proposed Endangered","Proposed Threatened","Species of Concern","Threatened","Under Review") & Main.Category == "Animal") %>%
  ggplot(aes(x=reorder(Category, Category, function(x) length(x)), fill=Conservation.Status)) +
  geom_bar() 
```

```{r}
library(reshape2)
df_bar <- df %>% filter(Main.Category == "Plant") %>% select(Conservation.Status, Category)
#dropping the column Park.Code
df_bar$Park.Code <- NULL
Names = c("Alage","Fungi","Nonvascular Plant","Vascular Plant")
data1 <- data.frame(df_bar$Category, df_bar$Category) 
data2 <- melt(data1)

```


```{r}
df %>%
group_by(Category, Conservation.Status) %>%
summarize(n = n()) %>%
ggplot(aes(x = n, y = meanHeight, fill = Company)) +
geom_bar(aes(x = AreaCode, y = meanHeight, fill = Company), stat = "identity", position
= position_dodge()) +
labs(x = "Area Code", y = "Average Employee Height
(cm)", title = "Area Code v/s Average Height as per Company") +
coord_cartesian(ylim = c(150,175))
```


```{r}
df_bar <- df %>% filter(Main.Category == "Plant") %>% select(Conservation.Status, Category)
df_bar$Park.Code <- NULL
df_bar
count(df_bar, vars=c("Conservation.Status","Category"))
```

```{r}
park_state <- df %>% group_by(State) %>% summarise(n=n()) %>% arrange(desc(n)) %>% ungroup(park_state)
park_state
```

```{r}
count_park <- df_parks %>% group_by(State) %>% summarise(n=n())
head(arrange(count_park, desc(n)), n=10)
```





